Time Series Bootstrapping For Solar Current

### 2021-12-15 deliverables/paper.Rmd
Removed bindwith for new scale

This project began with the search for a research paper
utilizing the process of Monte Carlo sampling or other 
statistcal computing methods. We landed on a few papers 
that did exatcly this, one titled, "SolarStat: Modeling 
Photovoltaic Sources through Stochastic Markov Processes". 
Instead of comparing our works of Monte Carlo simulations 
we were able to utilize a more recent topic discussed in class.
This of course was bootsrapping, a topic we didn't get too in
depth about; however, everything needed to know will be explained 
later on in the paper. Regarding our choice of articles, it was a
question of interest and importance. During the last decade the world
has seen a massive uptick in renewable energy pushes. With this 
new wave of renewables and the rather large misunderstanding of their
total power we decided it would be interesting to look into solar and 
what it was all about. The big question being asked regarding
solar power is whether or not it is worth heading the initial cost
of set up and general mantienance. Many websites inform users of 
just this, according to the Solar.com Learning Center, the average
solar panel net cost is $2,700 per 1.5 kilowatt-hour system (@solar.com). 
The biggest upside to solar panels is being able to slice down your 
energy bill, and eventually have it pay for itself. In order to 
calculate a solar panels net output experts recommend looking at three things. 
1. Determine your locations peak sun hours (average US around 4 hours/day)
2. Determine your energy compensation (average $0.03/kilowatt-hour (kWh))
3. Determine your solar panels net energy output
Once you have calculated these three numbers it is as easy as 
multiplying them together to receive your daily solar panel 
compensation (@energysage). As an example Energy Sage looked at 
California receiving 5 peak hours of sunlight a day with a single 
solar panel producing 290 watts/hour. Daily, this panel will produce 
around 1.5 kWh per day yielding the owner a rough compensation 
of $16.42 a year.

This paper looks at data from a 12 year period to determine the current 
output of a specified city. It can be used to determine 
on a month by month basis if a location produces a sufficient amount 
of current from a solar panel. Providing customers with an idea on 
whether or not they should make the investment that is solar.

@@

To translate raw irradiance data into meaningful current data, several steps
had to be taken to simulate the electrical workings of a real PV module. First,
adjustments had to be made to account for the fact that an imperfect amount of
irradiance actually reaches a PV module for capture. Then, effective irradiance
must be translated into output current and maximized with an attached power
processor. The process and accompanying assumptions are given below.

@@

```{r filter, include=FALSE}
HOUR_WINDOW <- 11:16
df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
filtered_current_hist <- ggplot() +
  geom_histogram(aes(x = df_high_irradiance$current_out, y = ..density..), binwidth = .0025) +
  theme_bw() +
  labs(
    title = "Distribution of Current Outputs", subtitle = "Restricted to high irradiance hours of the day",
    x = "Current Output", y = "Density"
  )
print(filtered_current_hist)
```

@@

The main goal of this paper is to provide point and interval
estimates for average (simulated) current output
based on solar irradiance.
Given that the sample size is large,
a simple average may have been sufficient.
However, we were interested in characterizing
the difference in performance between
standard bootstrapping methods and dependent bootstrapping methods.
The current output provides a good test case for this.
Both the preprocessed and postprocessed currents are not independent data;
values of either irradiance or current output
depend on surrounding values in time.
As such, standard methods for bootstrapping should be
less effective at estimating the true mean irradiance output.

The standard (*iid*) bootstrap procedure begins with
the assumption that our original sample consists of
independent and identically distributed draws from a distribution.
The univariate bootstrap then provides a distribution of bootstrap
estimates for a statistic,
$\hat{\theta}^{(1)}, \ldots, \{\hat{\theta}^{(B)}$,
where $B$ is the number of bootstrap replicates
and $\theta$ is a statistic of interest.
For our purposes, $\theta$ is simply the average current output
for our specific solar module configuration.
The *iid* bootstrap constructs the distribution of bootstrap estimates
as follows:

@@

In contrast, a dependent bootstrap does not assume
the original sample was independent and identically distributed.
To capture the dependence in the data,
we chose to use a nonparametric non-overlapping block bootstrap.
This is because during the data filtering step,
we excluded nighttime values from the dataset.
This meant we essentially had predefined,
mostly independent blocks from which to sample:
the daytime current values for each day.
Given that our daytime current values for each day
could be used as non-overlapping blocks,
our dataset could be seen as a collection
of $n_B$ blocks.

@@

time_series_ghi <- time_series_of_var(df, "GHI", "GHI", 6)
time_series_current <- time_series_of_var(df, "current_out", "Output current", 6)
ggsave("./images/time_series_ghi.png", make_plot_for_pres(time_series_ghi))
ggsave("./images/time_series_current.png", make_plot_for_pres(time_series_current))
```

@@

hour_plot_ghi <- hour_plot_of_var(df, "GHI", "GHI")
hour_plot_current <- hour_plot_of_var(df, "current_out", "output current")
ggsave("./images/hour_plot_ghi.png", make_plot_for_pres(hour_plot_ghi))
ggsave("./images/hour_plot_current.png", make_plot_for_pres(hour_plot_current))
```

@@

```{r current-ts, fig.cap="Time series of simulated current values for the first six days in the dataset.", echo=FALSE}
print(time_series_current)
```

@@

```{r current-hour-boxplot, fig.cap="Distribution of GHI at each hour for the entire unfiltered dataset.", echo=FALSE}
print(hour_plot_current)
```

@@

#' Given a dataframe, return a matrix of bootstrap blocks
#'
#' @param df A dataframe of irradiance values
#' @return A matrix whose rows are bootstrap blocks
make_blocks_from_df <- function(df) {
  df <- df %>%
    mutate(date = lubridate::make_date(Year, Month, Day))
  # The number of blocks is the number of days
  number_of_blocks <- length(unique(df$date))
  # The length of a block is the number of
  # observations in each individual day.
  # In this case, we can simply look at the first one,
  # because all of the days have the same number of observations
  length_of_block <- nrow(df[df$date == unique(df$date)[1], ])
  blocks <- matrix(NA, nrow = number_of_blocks, ncol = length_of_block)
  start <- 1
  stop <- length_of_block
  for (i in 1:number_of_blocks) {
    # blocks[i, ] <- df$current_out[(i - 1) * length_of_block + 1 : (i * length_of_block)]
    blocks[i, ] <- df$current_out[start:stop]
    start <- start + length_of_block
    stop <- stop + length_of_block
  }
  return(blocks)
}
# We have to use the modified DataFrame here,
# or else the assumptions about block size don't hold
test <- make_blocks_from_df(df_high_irradiance)

@@

iid_bootstrap_estimates_from_df <- function(df, B = 10000) {
  boot_current <- boot(df$current_out, meanv, B)
  return(boot_current$t)
}

@@

mean_boot_plot <- ggplot(vis_df) +
  geom_histogram(aes(x = boot, y = ..density.., fill = type),
    position = "identity", alpha = .5, color = "black",
    binwidth = .0001
  ) +
  # geom_segment(aes(x = block_ci[1], xend = block_ci[2], y = 1250, yend = 1250), color = 'tomato', alpha = .01) +
  geom_vline(aes(xintercept = block_ci[1]), linetype = "dashed", color = "tomato", alpha = .5) +
  geom_vline(aes(xintercept = block_ci[2]), linetype = "dashed", color = "tomato", alpha = .5) +
  # geom_segment(aes(x = iid_ci[1], xend = iid_ci[2], y = 1300, yend = 1300), color = 'turquoise3', alpha = .01) +
  geom_vline(aes(xintercept = iid_ci[1]), linetype = "dashed", color = "turquoise3", alpha = .5) +
  geom_vline(aes(xintercept = iid_ci[2]), linetype = "dashed", color = "turquoise3", alpha = .5) +
  labs(
    fill = "Bootstrap Type", title = "Bootstrapped Current Output", subtitle = "Percentile confidence intervals given",
    x = "Bootstrap Sample Means", y = "Density"
  ) +
  theme_bw()
ggsave("./images/mean_boot_plot.png", make_plot_for_pres(mean_boot_plot), width = 16, height = 12)
```

```{r mean-boot-plot, fig.cap="Distributions of bootstrap mean simulated current for both iid and block bootstrapping methods. Percentile 95 percent bootstrap confidence intervals for mean simulated current overlayed.", echo=FALSE}
print(mean_boot_plot)
```

@@

bootstrap_ci_table %>% kable(booktabs = TRUE,
caption="95% confidence intervals for the mean simulated current, as generated by each of the two bootstrapping methods.")

@@

sixhr_plot_current <- time_series_irradiance(df_high_irradiance, "current_out", "Current Output", 6)
ggsave("./images/sixhr_plot_current.png", make_plot_for_pres(sixhr_plot_current))

HOUR_WINDOW <- 8:19
df_more_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
filtered_current_hist <- ggplot() +
  geom_histogram(aes(x = df_high_irradiance$current_out)) +
  theme_bw()
print(filtered_current_hist)

twelvehr_plot_current <- time_series_irradiance(df_more_irradiance, "current_out", "Current Output", 12)
ggsave("./images/twelvehr_plot_current.png", make_plot_for_pres(twelvehr_plot_current))

day_plot_current <- time_series_irradiance(df, "current_out", "Current Output", 24)
ggsave("./images/day_plot_current.png", make_plot_for_pres(day_plot_current))

@@

For future work, it would be good to examine the wider implications
of the inferential results explored in this paper.
The model paper for this project simulated results for small solar panels
in a specific configuration,
making it hard to generalize the results to any meaningful solar array.
However, if the results could be generalized,
it would be good to see how current output based on solar irradiance
fits into a larger model of uncertainty surrounding renewable energy generation.
One example of such an approach would be @constante_2018,
in which PV output simulated from solar irradiance
is used in conjunction with other sources of of uncertainty.
The estimates of PV output in this paper are only part
of the story with respect to uncertainty in PV power generation.
Other relevant types of uncertainty are uncertainty of energy demand,
uncertainty in personal PV module placement on the grid,
and uncertainty in the power generation of other renewable energy
sources, to name a few (@constante_2018).

>>>>

This project began with the search for a research paper
utilizing the process of Monte Carlo sampling or other 
statistcal computing methods. We landed on a few papers 
that did exatcly this, one titled, "SolarStat: Modeling 
Photovoltaic Sources through Stochastic Markov Processes". 
Instead of comparing our works of Monte Carlo simulations 
we were able to utilize a more recent topic discussed in class.
This of course was bootsrapping, a topic we didn't get too in
depth about; however, everything needed to know will be explained 
later on in the paper. Regarding our choice of articles, it was a
question of interest and importance. During the last decade the world
has seen a massive uptick in renewable energy pushes. With this 
new wave of renewables and the rather large misunderstanding of their
total power we decided it would be interesting to look into solar and 
what it was all about. The big question being asked regarding
solar power is whether or not it is worth heading the initial cost
of set up and general mantienance. Many websites inform users of 
just this, according to the Solar.com Learning Center, the average
solar panel net cost is \$2,700 per 1.5 kilowatt-hour system (@solar.com). 
The biggest upside to solar panels is being able to slice down your 
energy bill, and eventually have it pay for itself. In order to 
calculate a solar panels net output experts recommend looking at three things. 
1. Determine your locations peak sun hours (average US around 4 hours/day)
2. Determine your energy compensation (average \$0.03/kilowatt-hour (kWh))
3. Determine your solar panels net energy output
Once you have calculated these three numbers it is as easy as 
multiplying them together to receive your daily solar panel 
compensation (@energysage). As an example Energy Sage looked at 
California receiving 5 peak hours of sunlight a day with a single 
solar panel producing 290 watts/hour. Daily, this panel will produce 
around 1.5 kWh per day yielding the owner a rough compensation 
of \$16.42 a year.

This paper looks at data from a 12 year period to determine the power 
output of a small grid of solar panels
located in Los Angeles, California,
based on solar radiation data.
The information gleaned from simulating power output
can determine if,
on an overall basis, a location produces a sufficient amount 
of power from a solar panel.
This gives customers information to determine
whether or not they should make an investment in solar energy.

@@

To translate raw irradiance data into meaningful power data, several steps
had to be taken to simulate the electrical workings of a real PV module. First,
adjustments had to be made to account for the fact that an imperfect amount of
irradiance actually reaches a PV module for capture. Then, effective irradiance
must be translated into output current and maximized with an attached power
processor. The process and accompanying assumptions are given below.

@@

```{r filter, include=FALSE}
HOUR_WINDOW <- 11:16
df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
filtered_power_hist <- ggplot() +
  geom_histogram(aes(x = df_high_irradiance$power_out, y = ..density..), binwidth = .0025) +
  theme_bw() +
  labs(
    title = "Distribution of Power Outputs", subtitle = "Restricted to high irradiance hours of the day",
    x = "Power Output", y = "Density"
  )
print(filtered_power_hist)
```

@@

The main goal of this paper is to provide point and interval
estimates for average (simulated) power output
based on solar irradiance.
Given that the sample size is large,
a simple average may have been sufficient.
However, we were interested in characterizing
the difference in performance between
standard bootstrapping methods and dependent bootstrapping methods.
The power output provides a good test case for this.
Both the preprocessed GHI
and postprocessed power outputs are not independent data;
values of either irradiance or power output
depend on surrounding values in time.
As such, standard methods for bootstrapping should be
less effective at estimating the true mean irradiance output.

The standard (*iid*) bootstrap procedure begins with
the assumption that our original sample consists of
independent and identically distributed draws from a distribution.
The univariate bootstrap then provides a distribution of bootstrap
estimates for a statistic,
$\hat{\theta}^{(1)}, \ldots, \{\hat{\theta}^{(B)}$,
where $B$ is the number of bootstrap replicates
and $\theta$ is a statistic of interest.
For our purposes, $\theta$ is simply the average power output
for our specific solar module configuration.
The *iid* bootstrap constructs the distribution of bootstrap estimates
as follows:

@@

In contrast, a dependent bootstrap does not assume
the original sample was independent and identically distributed.
To capture the dependence in the data,
we chose to use a nonparametric non-overlapping block bootstrap.
This is because during the data filtering step,
we excluded nighttime values from the dataset.
This meant we essentially had predefined,
mostly independent blocks from which to sample:
the daytime power values for each day.
Given that our daytime power values for each day
could be used as non-overlapping blocks,
our dataset could be seen as a collection
of $n_B$ blocks.

@@

time_series_ghi <- time_series_of_var(df, "GHI", "GHI", 6)
time_series_power <- time_series_of_var(df, "power_out", "Output power", 6)
ggsave("./images/time_series_ghi.png", make_plot_for_pres(time_series_ghi))
ggsave("./images/time_series_power.png", make_plot_for_pres(time_series_power))
```

@@

hour_plot_ghi <- hour_plot_of_var(df, "GHI", "GHI")
hour_plot_power <- hour_plot_of_var(df, "power_out", "Output Power")
ggsave("./images/hour_plot_ghi.png", make_plot_for_pres(hour_plot_ghi))
ggsave("./images/hour_plot_power.png", make_plot_for_pres(hour_plot_power))
```

@@

```{r power-ts, fig.cap="Time series of simulated power values for the first six days in the dataset.", echo=FALSE}
print(time_series_power)
```

@@

```{r power-hour-boxplot, fig.cap="Distribution of GHI at each hour for the entire unfiltered dataset.", echo=FALSE}
print(hour_plot_power)
```

@@

#' Given a dataframe, return a matrix of bootstrap blocks
#'
#' @param df A dataframe of irradiance values
#' @return A matrix whose rows are bootstrap blocks
make_blocks_from_df <- function(df) {
  df <- df %>%
    mutate(date = lubridate::make_date(Year, Month, Day))
  # The number of blocks is the number of days
  number_of_blocks <- length(unique(df$date))
  # The length of a block is the number of
  # observations in each individual day.
  # In this case, we can simply look at the first one,
  # because all of the days have the same number of observations
  length_of_block <- nrow(df[df$date == unique(df$date)[1], ])
  blocks <- matrix(NA, nrow = number_of_blocks, ncol = length_of_block)
  start <- 1
  stop <- length_of_block
  for (i in 1:number_of_blocks) {
    # blocks[i, ] <- df$current_out[(i - 1) * length_of_block + 1 : (i * length_of_block)]
    blocks[i, ] <- df$power_out[start:stop]
    start <- start + length_of_block
    stop <- stop + length_of_block
  }
  return(blocks)
}
# We have to use the modified DataFrame here,
# or else the assumptions about block size don't hold
test <- make_blocks_from_df(df_high_irradiance)

@@

iid_bootstrap_estimates_from_df <- function(df, B = 10000) {
  boot_power <- boot(df$power_out, meanv, B)
  return(boot_power$t)
}

@@

mean_boot_plot <- ggplot(vis_df) +
  geom_histogram(aes(x = boot, y = ..density.., fill = type),
    position = "identity", alpha = .5, color = "black"
  ) +
  # geom_segment(aes(x = block_ci[1], xend = block_ci[2], y = 1250, yend = 1250), color = 'tomato', alpha = .01) +
  geom_vline(aes(xintercept = block_ci[1]), linetype = "dashed", color = "tomato", alpha = .5) +
  geom_vline(aes(xintercept = block_ci[2]), linetype = "dashed", color = "tomato", alpha = .5) +
  # geom_segment(aes(x = iid_ci[1], xend = iid_ci[2], y = 1300, yend = 1300), color = 'turquoise3', alpha = .01) +
  geom_vline(aes(xintercept = iid_ci[1]), linetype = "dashed", color = "turquoise3", alpha = .5) +
  geom_vline(aes(xintercept = iid_ci[2]), linetype = "dashed", color = "turquoise3", alpha = .5) +
  labs(
    fill = "Bootstrap Type", title = "Bootstrapped Power Output", subtitle = "Percentile confidence intervals given",
    x = "Bootstrap Sample Means", y = "Density"
  ) +
  theme_bw()
ggsave("./images/mean_boot_plot.png", make_plot_for_pres(mean_boot_plot), width = 16, height = 12)
```

```{r mean-boot-plot, fig.cap="Distributions of bootstrap mean simulated power for both iid and block bootstrapping methods. Percentile 95 percent bootstrap confidence intervals for mean simulated power overlayed.", echo=FALSE}
print(mean_boot_plot)
```

@@

bootstrap_ci_table %>% kable(booktabs = TRUE,
caption="95% confidence intervals for the mean simulated power, as generated by each of the two bootstrapping methods.")

@@

sixhr_plot_power <- time_series_irradiance(df_high_irradiance, "power_out", "Power Output", 6)
ggsave("./images/sixhr_plot_power.png", make_plot_for_pres(sixhr_plot_power))

HOUR_WINDOW <- 8:19
df_more_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
filtered_power_hist <- ggplot() +
  geom_histogram(aes(x = df_high_irradiance$power_out)) +
  theme_bw()
print(filtered_power_hist)

twelvehr_plot_power <- time_series_irradiance(df_more_irradiance, "power_out", "Power Output", 12)
ggsave("./images/twelvehr_plot_power.png", make_plot_for_pres(twelvehr_plot_power))

day_plot_power <- time_series_irradiance(df, "power_out", "Power Output", 24)
ggsave("./images/day_plot_power.png", make_plot_for_pres(day_plot_power))

@@

For future work, it would be good to examine the wider implications
of the inferential results explored in this paper.
The model paper for this project simulated results for small solar panels
in a specific configuration,
making it hard to generalize the results to any meaningful solar array.
However, if the results could be generalized,
it would be good to see how power output based on solar irradiance
fits into a larger model of uncertainty surrounding renewable energy generation.
One example of such an approach would be @constante_2018,
in which PV output simulated from solar irradiance
is used in conjunction with other sources of of uncertainty.
The estimates of PV output in this paper are only part
of the story with respect to uncertainty in PV power generation.
Other relevant types of uncertainty are uncertainty of energy demand,
uncertainty in personal PV module placement on the grid,
and uncertainty in the power generation of other renewable energy
sources, to name a few (@constante_2018).
